---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# bSantiago

<!-- badges: start -->
<!-- badges: end -->

bSantiago is a data package with the results of a 2016 survey conducted in Santiago, Chile, designed to study the health and psychological status of residents in the city with respect to their daily commuting/travel experience. The survey provides data to examine feelings and values with respect to active and motorized modes of transportation.

## Installation

You can install the development version of Santiago from [GitHub](https://github.com/) with:

``` r
install.packages("remotes")
remotes::install_github("paezha/bSantiago")
```

## Example

This example shows how to load one of the data tables included in the package:

```{r introduction}
library(bSantiago)
data("Santiago_BE")
```

The data table `Santiago_BE` includes information about the built environment around the place of residence of respondents:
```{r summary}
summary(Santiago_BE)
```
