---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

```{r flowchart ,include=FALSE}
library(DiagrammeR)

DiagrammeR::grViz("digraph {

graph [layout = dot, rankdir = TB]

# define the global styles of the nodes. We can override these in box if we wish
node [shape = rectangle, style = filled, fillcolor = Linen]

data1 [label = 'Survey Data Preprocessing Guide', shape = rectangle, fillcolor = Beige]

1 [label =  'Navigation to Data Preprocessing Script \n  from data-raw/00-Data-Preprocessing.Rmd']


2 [label = 'Package Installation \n and Loading']
3 [label = 'Loading Communa \n (Municipality) Data
']
4 [label = 'Reading and Translating \n Survey Data']
5 [label = 'Table separation based on \n subject and lable conversion']
6 [label = 'Saving Data Frames']

7 [label = 'Steps for splitting multiple answers  \n in a single cell']


# edge definitions with the node IDs
{data1}  -> 1 -> 2 -> 3-> 4 -> 5 -> 6-> 7
}")


```




# bSantiago
## Introduction

In this notebook we read and preprocess the raw data collected as a paper-based survey conducted face-to-face in Santiago in 2016 by Dr. Beatriz Mella-Lira. The survey collected information on a wide range of travel-related issues (socio-demographics, health-related, perceptions and travel behaviour, travel choices and planning, social interaction factors, built environment, and so on). 

The data collection considered a quota-sampling method based on the information from the Pre-census of 2012, and with a total of 451 persons. 

## Installation

You can install the development version of Santiago from [GitHub](https://github.com/) with:

``` r
install.packages("remotes")
remotes::install_github("paezha/bSantiago")
```

## Example

This example shows how to load one of the data tables included in the package:

```{r introduction}
library(bSantiago)
data("Santiago_BE")
```

The data table `Santiago_BE` includes information about the built environment around the place of residence of respondents:
```{r summary}
summary(Santiago_BE)
```
