---
title: "A dataset to study equity in individual travel behavior and choices in Santiago, Chile"
author: "Niloofar Nalaee, Beatriz Mella Lira, Antonio Paez"
format: docx
---

```{r}
#| label: load-packages
#| echo: false
library(bSantiago) # A data package with the results of a travel and well-being survey conducted in Santiago in 2016
library(skimr) # Compact and Flexible Summaries of Data
```

# Abstract

The dataset presented in this article would be advantageous to give intuition about the experience of daily travel and its psychological impact on travelers, ranging from positive feelings of enjoyment in some to the sensation of stress in many others. This dataset particularly examines the feelings of stress by users of active and motorized modes of transportation. Furthermore, it also investigates the importance that travelers attach to their feelings of stress. This allows us to explore the concept of “limited horizons”, the normalization of subpar experiences by those less able to adapt. This data provides information about commuters in terms of their individual characteristics, health-related factors while using transportation modes, feelings and emotions towards different modes of travel, own decisions, social interaction, attitudes towards nature and sustainability, shifts between modes, attitudes towards built environment, traveling to work. Data for the research are drawn from a survey conducted in Santiago, Chile, based on a quota-sampling method based on the information from Pre-Census of 2012, and in total, 451 persons validly completed the survey. Describe the statistics and conclude suggest that

## Key words 

## Specifications Table

See @tbl-specifications for details of the data set.

| Items | Explanation |
|:--------|:-----|
| Subject area      | Transportation, Geography, Public Health and Health Policy, Urban development   |
| More specific subject area     | Transport inequalities, Stress and limited horizons, Travel behavior, Global South  |
| Type of data       | R Data Package    |
| How data was acquired | The survey conducted using questionnaire. The instrument contains descriptive data of respondents and 5-Likert scale questionnaire regarding most sections of the questionnaire|
| Data format | Raw Data | 
| Parameters for data collection | The survey collected by a quota-sampling method based on the information from Pre-Census of 2012, and in total, 451 persons validly completed the survey and face-to-face in Santiago, Chile in 2016. The survey collected information on a wide range of travel-related issues (socio-demographics, health-related, perceptions and travel behavior, travel choices and planning, social interaction factors, built environment, among others) |
| Description of data collection | Data was acquired through the 5-Likert scale questionnaire regarding most sections of the questionnaire, using a face-to-face and quota-sampling method for individual characteristics  | 
| Data source location | Santiago, Chile | 
| Data accessibility | State if data is with this article or in public repository.  If public repository, please explicitly name repository and data identification number and provide a direct URL to data | 

: Specifications table {#tbl-specifications}

## Value of the data 

A large body of the data has made inroads investigating psychological impact on travelers ranging from positive feelings of enjoyment in some to the sensation of stress in many others that can affect the effectiveness of policy measure (in the case of positive feelings) and are known to affect health outcomes (in the case of stress). This would be interesting for those with transport policies concerns.
Dataset contribute to psychological impact on travelers both active and motorized modes of transportation to examines not only the feeling of stress, but also how these effects are experienced by travelers and investigates the importance that travelers attach to their feelings of stress which makes it valuable for researchers who focused on public sector development and health-related policies.
This dataset allows us to explore the concept of “limited horizons”, the normalization of subpar experiences by those less able to adapt, an advantageous resource for further research regarding transport inequalities, index of stress, travel behavior in the region or even as a representative for other areas with similar attributes.
The dataset provides a wide range of travel-related issues such as socio-demographics, health-related, perceptions and travel behavior, travel choices and planning, social interaction factors, built environment, among others.  

## Data
 
The dataset described in this paper contains individual characteristics, health information, feelings and emotions information, reason and planning decision of respondents, social interaction of respondents, nature and sustainability, telecommunication and shifting, built environment and travel-work related information. As can be seen in data set, we have a wide variety of variables in each section.
Data collected regarding individual characteristics reveals interesting difference when it comes to gender, age, income, main mode of transport and so on. Among persons who responded to the survey, males have far more higher income levels than females. It can be seen among people who have a salary more than 997 dollars and 18 aged until 54, using different modes of travel as the main mode is almost similar in terms of using car and metro.These information give insights about different segments of gender, age and so on to organize planning decisions for them according to their requirements. Also this part would be advantageous for business to figuring out marketing and applying strategies for future patterns based on consumer demands.

## Running Code

```{r Individual characteristics}
#| label: Individual-characteristics-Descriptives
#| echo: false
skim(Santiago_IC)
```


Health information could be useful in investigating affects of transportation policy decisions on public health.
```{r Health Information}
#| label: Health-Information-Descriptives
#| echo: false
skim(Santiago_H)
```

```{r Feelings and Emotions}
#| label: Feelings-Emotions-Descriptives
#| echo: false
skim(Santiago_FE)
```

```{r Reason and Planning Decision}
#| label: Reason-Planning-Decision-Descriptives
#| echo: false
skim(Santiago_RPD)
```

```{r Social Interaction}
#| label: Social-Interaction-Descriptives
#| echo: false
skim(Santiago_SI)
```

```{r Nature and Sustainability}
#| label: Nature-Nustainability-Descriptives
#| echo: false
skim(Santiago_NS)
```

```{r ITC and Shifting}
#| label: ITC-Shifting-Descriptives
#| echo: false
skim(Santiago_IS)
```

```{r Built Environment}
#| label: Built-Environment-Descriptives
#| echo: false
skim(Santiago_BE)
```
## Experimental Design, Materials and Methods

The study is based on a paper-based survey conducted face-to-face in Santiago in 2016. The survey collected information on a wide range of travel-related issues (socio-demographics, health-related, perceptions and travel behavior, travel choices and planning, social interaction factors, built environment, among others). The data collection considered a quota-sampling method based on the information from Pre-Census of 2012, and in total, 451 persons validly completed the survey. This paper considers the first part of the survey, with information about the basic socio-economic data, travel choices, activities and commuting information, and the question related to the levels of stress experienced in while traveling.

## Acknowledgements

## References 
